### 2015-03-04 grassposter/2015_EGU_Python/listings/grass_module_cli.tex
python poster: blocks for outline, bash in documentation to python block, generated interface block, width according to other EGU posters
git-svn-id: https://svn.osgeo.org/grass/grass-promo@64799 15284696-431f-4ddb-bdfa-cd5b030d7da7

% taken from grass-promo/grassposter, written by me I think

\documentclass{standalone}

\usepackage[active,tightpage]{preview}
\usepackage{listings}
\PreviewEnvironment{lstlisting}

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\input{listings_settings}

\begin{document}

% do some trim here, do more trim when including if needed
\begin{lstlisting}[linewidth=34.2em]
Description:
 Adds the values of two rasters (A + B)

Keywords:
 raster, algebra, sum

Usage:
 r.plus araster=name braster=name output=name [--overwrite] [--help]
   [--verbose] [--quiet] [--ui]

Flags:
 --o   Allow output files to overwrite existing files
 --h   Print usage summary
 --v   Verbose module output
 --q   Quiet module output
 --ui  Force launching GUI dialog

Parameters:
  araster   Name of input raster A in an expression A + B
  braster   Name of input raster B in an expression A + B
   output   Name for output raster map
\end{lstlisting}
\end{document}