### 2015-03-26 grassposter/2015_EGU_G7_Python/listings/tgrass_examples.tex
Fixed wrong raster object open method
git-svn-id: https://svn.osgeo.org/grass/grass-promo@64924 15284696-431f-4ddb-bdfa-cd5b030d7da7

% taken from grass-promo/grassposter, written by me I think

\documentclass{standalone}

\usepackage[active,tightpage]{preview}
\usepackage{listings}
\PreviewEnvironment{lstlisting}

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\input{listings_settings}

\begin{document}

\begin{lstlisting}[language=Python,linewidth=34.2em]
# Import and initialize the GRASS GIS temporal framework
import grass.temporal as tgis
tgis.init()
# Open an existing space time raster dataset (STRDS)
temp_strds = tgis.open_old_stds(name="temp_daily", 
                                type="strds")
# Shift all registered raster map layer 2 days in the future
temp_strds.shift("2 days")
# Open an exitsing time stamped raster map as map object
temp_24Mar77 = tgis.RasterDataset('temp_1977_Mar_24@soeren')
# Unregister the raster map from the STRDS
temp_strds.unregister_map(temp_24Mar77)
# Get a list of all registered raster map layers with
# a start time later 1990-05-22 from the STRDS
maps = temp_strds.get_registered_maps_as_objects(\
                     where="start_time > '1990-05-22'")
# Create a temporal buffer of 6h for each map in the list
for map in maps:
    map.temporal_buffer("6 hours", update=True)
# Compute new spatio-temporal extent and granularity 
temp_strds.update_from_registered_maps()
# Get the granuarity of the STRDS
gran = temp_strds.get_granularity()
\end{lstlisting}
\end{document}