### 2015-04-01 grassposter/2015_EGU_G7_Python/listings/pygrass_raster_examples.tex
rewrite Highlights, parser box, IPython box, revert temporal box to https://trac.osgeo.org/grass/changeset/64917 and other improvements
git-svn-id: https://svn.osgeo.org/grass/grass-promo@64967 15284696-431f-4ddb-bdfa-cd5b030d7da7

% taken from grass-promo/grassposter, written by me I think

\documentclass{standalone}

\usepackage[active,tightpage]{preview}
\usepackage{listings}
\PreviewEnvironment{lstlisting}

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\input{listings_settings}

\begin{document}

\begin{lstlisting}[language=Python,linewidth=44.2em]
import numpy as np
from grass.pygrass.raster import RasterRow
from grass.pygrass.raster.buffer import Buffer
from grass.pygrass.gis.region import Region

def raster2numpy(name, mapset):
    """Return a numpy array from a raster map"""
    with RasterRow(name, mapset, mode='r') as array:
        return np.array(array)

def numpy2raster(array, mtype, name):
    """Save a numpy array to a raster map"""
    reg = Region()
    if (reg.rows, reg.cols) != array.shape:
        msg = "Region and array shape are different: %r != %r"
        raise TypeError(msg % ((reg.rows, reg.cols), array.shape))
    with RasterRow(name, mode='w', mtype=mtype) as new:
        newrow = Buffer((array.shape[1],), mtype=mtype)
        for row in array:
            newrow[:] = row[:]  # cast array to raster type
            new.put_row(newrow)  # write row to raster map
# note that there is a specialized package which implements this functionality
\end{lstlisting}
\end{document}