### 2018-11-29 grassposter/2018_AGU_Geospatial_Research_Tool/poster.tex
poster: shorten, increase font size, set svn props
git-svn-id: https://svn.osgeo.org/grass/grass-promo@73725 15284696-431f-4ddb-bdfa-cd5b030d7da7

% License: CC BY-SA
% Authors: See the authors below and see also acknowledgement for authors of some images or research

\documentclass[25pt, margin=0mm, innermargin=25mm, blockverticalspace=25mm, colspace=25mm, subcolspace=8mm]{tikzposter}
\geometry{paperwidth=63in,paperheight=42in}

% to stretch boxes over whole paper with custor paper size
\makeatletter
\setlength{\TP@visibletextwidth}{\textwidth-2\TP@innermargin}
\setlength{\TP@visibletextheight}{\textheight-2\TP@innermargin}
\makeatother

% Fira Sans, GRASS GIS branding sans serif font
\usepackage{FiraSans}
\renewcommand*\oldstylenums[1]{{\firaoldstyle #1}}

% EB Garamond, GRASS GIS branding serif font
% note that EB Garamond does not have bold
\usepackage[cmintegrals,cmbraces]{newtxmath}
\usepackage{ebgaramond-maths}

% might be needed for both font packages
\usepackage[T1]{fontenc}

% uncomment for all sans serif
% \renewcommand{\familydefault}{\sfdefault}

\usepackage[utf8]{inputenc}
\usepackage{wrapfig}
\usepackage[hidelinks]{hyperref}

% For bibliography styling
%% TODO: all names should be abbreviated
\usepackage{natbib}

\definecolor{textcolor}{HTML}{000000}

\definecolor{titleTextColor}{HTML}{000000}
\definecolorpalette{grassColorPalette} {
  \definecolor{colorOne}{HTML}{419041}
  % \definecolor{colorTwo}{HTML}{cccccc}
  \definecolor{colorTwo}{HTML}{dddddd}
  \definecolor{colorThree}{HTML}{F1B52D}
  % \definecolor{colorThree}{HTML}{EFA126}
}

\usetheme{Default}
\usetitlestyle{Empty}
\usecolorstyle[colorPalette=grassColorPalette]{Britain}
\colorlet{backgroundcolor}{white}

\title{
\Huge
\textcolor{titleTextColor}{
\textsf{
% \textbf{
\fontsize{130}{100}\selectfont
\textbf{GRASS}\,{\firalight GIS}: A General-purpose Geospatial Research Tool
% }
}
}
}

\newlength{\grasslogoheight}
\setlength{\grasslogoheight}{0.09\textheight}
\newlength{\instlogoheight}
\setlength{\instlogoheight}{0.33\grasslogoheight}

% \setlength{\blocktitleheight}{0.02\textheight}

% style for institute numbers
\newcommand{\inst}[1]{\hspace{2pt}$^{\mbox{\normalsize#1}}$\hspace{-7pt}}
\newcommand{\instlist}[1]{\hspace{1pt}$^{\mbox{\normalsize#1}}$\hspace{2pt}}

\author{
Helena Mitasova\inst{1},
Vaclav Petras\inst{1}\,*\hspace{-7pt},
Anna Petrasova\inst{1},
\&
Markus Neteler\inst{2}
\\
% AGU allows to specify a scientific team, but it does not seem to fit
% with our case 100%; simply using formatting which is in the AGU program
% (but adding note)
Scientific Team: GRASS GIS Development Team**
}
\institute{
\large
\instlist{1}Center for Geospatial Analytics, North Carolina State University, USA;
\instlist{2}mundialis GmbH \& Co. KG, Germany;
*Corresponding author: wenzeslaus@gmail.com, vpetras@ncsu.edu;
% using Black Duck Open Hub to ``guestimate'' that
**Includes over 10 other members of the core team and numerous other contributors
\\[1.7cm]
\includegraphics[height=3cm]{ncstate}%
\hspace{1cm}%
\includegraphics[height=3cm]{cga}%
\hspace{3cm}%
\includegraphics[height=3.4cm]{mundialis}%
}

\hypersetup
{
    pdfauthor={H. Mitasova, V. Petras, A. Petrasova, M. Neteler},
    pdfsubject={AGU Fall Meeting 2018 Poster},
    pdftitle={GRASS GIS: A General-purpose Geospatial Research Tool},
    pdfkeywords={GIS, algorithms, methods, preservation, science, reproducibility}
}

% \usetemplate{1}
% \setinstituteshift{1}

% \setblocktitleheight{2}
% \setblockspacing{1}

\graphicspath{{images/}{logos/}}

\newcommand{\blocktitlewrap}[1]{\textnormal{\textsf{\textsc{\huge#1}}}}
% it is not possible (?) to change block title in the class, using wrapper
% the command introduced using:
%   sed -i 's/\\block{\([^}]*\)}/\\block{\\blocktitlewrap{\1}}/g' main.tex

\newcommand{\CustomBlockFontSize}{\LARGE}

% bullet point style
\renewcommand{\labelitemi}{\textcolor{gray}{$\bullet$}\hspace{0.5ex}}

% GRASS module
\newcommand{\gmodule}[1]{\href{http://grass.osgeo.org/grass74/manuals/#1.html}{\emph{#1}}}
\newcommand{\gamodule}[1]{\href{http://grass.osgeo.org/grass74/manuals/addons/#1.html}{\emph{#1}}}
\newcommand{\gmodulenolink}[1]{\emph{#1}}

\begin{document}

\node[above left,opacity=0.99,inner sep=0pt,outer sep=4cm] at (bottomleft -| topright)%
  {\includegraphics[width=0.2\paperwidth]{grass}};

\maketitle[width=0.92\textwidth]

% \maketitle
% \addlogo[north west]{(2,-1)}{9cm}{images/Grass_GIS}
%Please insert your institution logo here
% \addlogo[north east]{(-2,-2.5)}{4cm}{images/logo_FEM_CRI}
% \addlogo[north east]{(-2,-5.5)}{4cm}{images/NC_State_Seal}
% \addlogo[north east]{(-8,-2.5)}{4cm}{images/Logo_cvut}
% \addlogo[north east]{(-8,-6.5)}{4cm}{images/IWMI_logo}
% \addlogo[north east]{(-2,-10.5)}{4cm}{images/logo_ec-jrc}

\begin{columns}

% Abstract
%
% GRASS GIS (grass.osgeo.org) is an open source software for geospatial analysis,
% remote sensing, general geoprocessing and visualization. It is a
% community-driven project with over 35 years of continuous software development
% based on scientific expertise from many geospatial fields. It is characterized
% by long-term releases, stable APIs, and emphasis on science. GRASS GIS
% distribution strives to provide single integrated environment for 2D and 3D
% raster analysis, image processing, vector data analysis, and spatio-temporal
% data processing. It supports large raster files (billions of cells), vector
% topology, coupling with databases, and 64 bit memory. New code based on recent
% research is typically contributed to GRASS GIS Addons repository. Mature, widely
% used code is then moved to the main code base to maximize integration and
% availability. Whether it is addons or the main code base, code is usually
% maintained by the community and preserved in long term even in cases when the
% original author no longer supports the code. To support the needs of scientists,
% the documentation includes not only links to the source code, its history, and
% its authors, but also links to research papers that describe the algorithms
% implemented in the modules.
%
% Code is not only maintained but also extended and improved. For example,
% watershed and stream extraction using least cost path approach was implemented
% in 1989 and extended for massive datasets in 2011. Similarly, vector topology
% cleaning was introduced in 2002, updated over time and substantially improved in
% 2016. Another example is a solar energy module available since 1993 and
% parallelized in 2017. The current stable version of GRASS GIS 7.4 provides new
% features for geosciences such as temporal framework with temporal algebra for
% large time-series processing, fusion of elevation models from various sensors,
% 3D flows, landform detection, image segmentation, simplified batch processing,
% and integrated Python editor among others. GRASS GIS runs in various
% environments including Linux, Mac, Windows, Docker, Raspberry Pi, and on HPC
% clusters. The modules are written in Python, C, or C++. Besides command line
% interface, GRASS GIS provides a graphical user interface, a Python API, and a C
% API. Interfaces with other languages such as R and Ruby are supplied by
% collaborating communities.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\column{0.25}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Overview and Functionality}}{

\LARGE

\begin{itemize}
% Basic Topics
 \item Community-driven project
 \item 35 years of continuous software development
 \item long-term releases, stable APIs, and emphasis on science
 \item single integrated environment for 2D and 3D raster analysis, image processing, vector data analysis, and spatio-temporal data processing
% GRASS GIS
% distribution strives to provide single integrated environment for 2D and 3D
% raster analysis, image processing, vector data analysis, and spatio-temporal
% data processing. It supports large raster files (billions of cells), vector
% topology, coupling with databases, and 64 bit memory.
%  \item To support the needs of scientists, the documentation includes not only links to the source code, its history, and
% its authors, but also links to research papers that describe the algorithms
% implemented in the modules.
\end{itemize}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Code Life Cycle}}{

\CustomBlockFontSize

\begin{itemize}
 \item New code based on recent research is typically contributed to GRASS GIS Addons repository.
 \item Mature, widely used code is then moved to the main code base to maximize integration and availability.
 \item Both main code base and Addons maintained by the community.
 \item Code preserved in long term even in cases when the original author no longer supports the code.
\end{itemize}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Example: Spatio-Temporal Data Analysis}}{

\CustomBlockFontSize

\begin{itemize}
 \item The time dimension was introduced in version 7.0 \citep{Gebbert20141, gebbert2015grass}
 \item Time series accessed as space time datasets and as individual vectors or rasters.
 \item More than 50 modules available to manage, analyze, process, and visualize space time datasets.
 \item More than 100,000 map layers can be now handled efficiently in GRASS GIS.
 \item Used for analysis of the European Climate Assessment \& Dataset ECA\&D \citep{Haylock2008_climate_series}
       and temperate climate zone identification \citep{Gebbert20141}.
 \item New temporal modules (e.g. \gmodule{t.rast.aggregate}) work beside well established \gmodule{r.series} module
       and specialized modules such as \gamodule{r.hants} implemented according to \cite{roerink2000reconstructing} or \gamodule{r.seasons}.
 \item Raster and vector temporal algebra can be used for tasks
       such as computing hydrothermal coefficient for a time series of climate data using the actual mathematical formula
       \citep{leppelt2015grass}.
\end{itemize}

\vspace*{1.5cm}

\begin{minipage}{\linewidth}
\centering
\includegraphics[width=.7\linewidth]{images/temporal_precip_temp}
\\
Creating a synchronized animation of monthly total precipitation and mean temperature for NC, USA
\end{minipage}

\vspace*{1cm}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\column{0.25}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Example: Water, Floods and Erosion}}{

\CustomBlockFontSize

\begin{itemize}
 \item The \gmodule{r.sim.water} simulation model \citep{Mitas1998b}
for overland flow with spatially variable rainfall excess conditions was integrated into the Emergency
Routing Decision Planning system as a WPS \citep{raghavan2014deploying}.
% TODO: paralelized
\item A unique least cost path algorithm which doesn't need any depression filling
was implemented in \gmodule{r.watershed} module in 1989
and it was updated for massive data sets in 2011.
% TODO: refs
\item The \gmodule{r.sim.water} module was also utilized by \cite{Petrasova2014} and is now part of
\emph{Tangible Landscape}, a tangible GIS system, which also incorporated \gmodule{r.damflood},
a dam break inundation simulation by \cite{cannata2012two}.
\end{itemize}

\bigskip

\vspace*{1cm}

\centering

\begin{minipage}{0.485\linewidth}
\centering
\includegraphics[width=0.7\textwidth]{rsimwater_architects}
\end{minipage}
~
\begin{minipage}{0.485\linewidth}
\includegraphics[width=\textwidth]{damflood_tangible}
\end{minipage}

\bigskip

\begin{minipage}{0.485\linewidth}
\centering
Overland flow simulated by \gmodule{r.sim.water} used for landscape
architecture design in Tangible Landscape environment
(Historical master plan for Lake Raleigh, NC, USA)
\end{minipage}
~
\begin{minipage}{0.49\linewidth}
Dam breach on Lake Raleigh (NC, USA) in Tangible Landscape environment simulated using \gamodule{r.damflood} module
\end{minipage}

% \vspace*{1.4cm}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Example: Vector Network Analysis}}{

\CustomBlockFontSize

\begin{itemize}
 \item Radim Blazek introduced several modules for vector network analysis in 2003
       including shortest path analysis (\gmodule{v.net.path}) and traveling salesman (\gmodule{v.net.salesman}).
 \item The number of available modules increased to almost 20.
 \item Stepan Turek implemented turns support into all relevant vector network modules in 2014.
%  \item Since version 7.0 fine control over node costs available.
% combination with r.cost/r.walk workflows \citep{Petrasova2014}
\end{itemize}

\bigskip

\centering
\begin{minipage}{0.9\linewidth}
\begin{center}
\includegraphics[width=\textwidth]{network_trips}
A series of trips to collect samples across North Carolina, USA
for water quality measurements
as part of the Polyfluorinated Alkyl Substance Testing (PFAST) project.
Trips are distinguished by color and were computed using multiple
iterations of traveling salesman (\gmodule{v.net.salesman}).
\end{center}
\end{minipage}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\column{0.25}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Example: Image Segmentations}}{

\CustomBlockFontSize

\begin{itemize}
 \item \gamodule{r.smooth.seg} was created by Alfonso Vitti \citep{vitti2008free, vitti2012mumford} and implemented a piece-wise smooth approximation of the original image which can be used to reduce noise in the original image.
% according to \cite{mumford1989optimal} and \cite{march1997variational}
 \item \gmodule{r.clump} for grouping pixels with the same integer values available since 1980s.
       The latest version of \gmodule{r.clump} supports multiple image bands and clumping of floating point values.
 \item Eric Momsen implemented region-growing image segmentation in 2012.
       Markus Metz then extended and optimized the code resulting in the inclusion of the \gmodule{i.segment} module into version 7.0.
       This replaced a multi-resolution classification module of the same name from 1992 \citep{zhuang1992image}.
 \item \gamodule{i.segment.hierarchical} by Pietro Zambelli is based on \gmodule{i.segment} and performs parallelized hierarchical segmentation.
 \item Rashad Kanavath and Markus Metz implemented SLIC Superpixels segmentation \citep{achanta2010epfl, achanta2012slic} in 2016 as \gmodule{i.superpixels.slic}.
\end{itemize}

\vspace*{0.7cm}

\begin{minipage}{0.5\linewidth}
\begin{center}
\includegraphics[width=\textwidth]{superpixels_slic_pseudo}
\end{center}
\end{minipage}
~
\begin{minipage}{0.5\linewidth}
\begin{center}
\includegraphics[width=\textwidth]{superpixels_slic_colored}
\end{center}
\end{minipage}
\vspace{2mm}
\begin{center}
Superpixels (black outlines) with pseudo-color and NDVI images, central Wake county, NC, USA
\end{center}

\vspace*{0cm}

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{References}}{

\vspace{-0.2cm}

\scriptsize

% \newcommand{\blocksectiontitle}[1]{\subsubsection*{\textcolor{gray}{\textsf{#1}}}}
\newcommand{\blocksectiontitle}[1]{\textbf{#1}}

%\blocksectiontitle{References}
\begingroup
\renewcommand{\section}[2]{}%
\bibliographystyle{apalike}
\bibliography{poster}
\endgroup

}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\column{0.25}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\block{\blocktitlewrap{Example: Lidar Data Processing}}{

\CustomBlockFontSize

\begin{itemize}
 \item Filtering of ground and non-ground points was included
early on in the lidar era in the \gmodule{v.lidar.edgedetection} group of modules.
% TODO: ref
% v.lidar.correction Corrects the v.lidar.growing output. It is the last of the three algorithms for LIDAR .
% v.lidar.edgedetection Detects the object's edges from a LIDAR data set.
% v.lidar.growing Building contour determination and Region Growing algorithm for determining the building inside
 \item The module \gmodule{v.surf.rst} for spatial interpolation was developed approximately 20 years
ago, but since then it has been improved several times \citep{tracvsurfrst}
including recent parallelization which will be included in GRASS GIS 7.4.
% TODO: ref for paralelized
 \item The module \gmodule{r.in.lidar} statistically analyzes massive point clouds.
% TODO: base_raster, ref
 \item The \gmodule{v.outlier} module serves as a base for a module \gamodule{v.lidar.mcc} implementing Multiscale Curvature Classification.
% TODO: ref
\end{itemize}

% v.delaunay v.voronoi New option to create Voronoi diagrams for areas.

\centering
\begin{minipage}{0.48\linewidth}
\centering
\includegraphics[width=0.7\textwidth]{elevation_lidar}
\end{minipage}
~
\begin{minipage}{0.48\linewidth}
\centering
\includegraphics[width=\textwidth]{lidar_profile}
\end{minipage}

\bigskip

\begin{minipage}{0.48\linewidth}
\centering
Digital elevation model interpolated from lidar data shows tillage in an field in Raleigh, NC, USA
\end{minipage}
~
\begin{minipage}{0.48\linewidth}
\centering
Vertical slice of a point cloud created with \gamodule{v.profile.points} shows tree structure in lidar data
\end{minipage}

}

\end{columns}

\end{document}